### 2018-11-30 Revision/Paper/method.tex
..

In this section, we present an efficient framework for computing the
active subspace. The framework is employed to the H$_2$/O$_2$ reaction kinetics
problem whereby the \rebut{logarithm of the} 
pre-exponent ($A_i$) in the rate law associated with individual reactions provided in
Table~\ref{tab:kinetics} is considered to be uniformly
distributed in the interval, \rebut{$[0.97\log(A_i^\ast), 1.03\log(A_i^\ast)]$};
$A_i^\ast$ is the nominal
estimate provided in~\cite{Yetter:1991}.
Two approaches are explored for estimating the gradient of ignition delay with
respect to \rebut{$\log(A_i)$}: 
a perturbation-based approach that involves computation of model
gradients using finite difference in order to construct the matrix $\hat{\mat{C}}$
in~\eqref{eq:chat}, and a regression-based approach 
that involves a linear regression fit to the available set of
model evaluations in order to approximate the gradient. 
The active subspace is computed in an iterative manner to avoid
unnecessary model evaluations once converged is established.

@@

As discussed earlier, gradient estimation using finite differences 
requires additional model evaluations at the
neighboring points in the input domain. 
Hence, for $N$ samples in a $d$-dimensional parameter space, $N(d+1)$
model evaluations are needed. On the other hand, gradient estimation
using the regression-based approach involves a series of linear regression
fits to subsets of available evaluations as discussed 
in~\cite{Constantine:2015} (see Algorithm 1.2). Hence, the computational effort is  
reduced by a factor $(d+1)$ when using the regression-based approach.
The specific sequence of steps
for computing the active subspace is discussed as follows.

We begin by evaluating the gradient of the model output, $\nabla_{\bm{\xi}}f$,
at an initial set of $n_0$ samples (generated using Monte Carlo sampling)
denoted by $\bm{\xi}_i$, $i$ = $1,\ldots,n_0$.
Using the gradient
evaluations, the matrix, $\hat{\mat{C}}$ is computed. Eigenvalue decomposition
of $\hat{\mat{C}}$ yields an initial estimate of the dominant eigenspace,
$\hat{\mat{W}}_1$ and the set of corresponding eigenvalues, 
$\hat{\mat{\Lambda}}_1$. Note that $\hat{\mat{W}}_1$ is obtained by
partitioning the eigenspace around $\lambda_j$ such that the ratio of
subsequent eigenvalues,
\rebut{$\left(\frac{\lambda_j}{\lambda_{j+1}}\right)\ge\mathcal{O}(10^1)$}.
 At each subsequent iteration, model evaluations are
generated at a new set of $n_k$ samples. The new set of gradient evaluations
are augmented with the available set to re-construct $\hat{\mat{C}}$ followed
by its eigenvalue decomposition. The relative change in the norm of the
difference in squared value of individual components of the dominant
eigenvectors between subsequent iterations is evaluated. The process is
terminated and the resulting eigenspace is considered to have converged once
the maximum relative change at iteration $k$, $\max(\delta \hat{\mat{W}}_{1,j}^{(k)})$
($j$ is used as an index for the eigenvectors),
is smaller than a given tolerance, $\tau$.  A regression fit to
$G(\hat{\mat{W}}_1^\top\bm{\xi})$ is used as a surrogate to characterize and
quantify the uncertainty in the model output. Moreover, the components of the
dominant eigenvectors are used to compute the activity scores, $\bm{\nu}_r(f)$,
which provide an insight into the relative importance of the uncertain inputs.
Note that the index, $r$, corresponds to the number of eigenvectors in
$\hat{\mat{W}}_1$. The sequence of steps as discussed are outlined
in Algorithm~\ref{alg:grad}.

@@

To assess its feasibility and suitability, we implement
Algorithm~\ref{alg:grad} to compute the active subspace for 
the 19-dimensional H$_2$/O$_2$ reaction kinetics
problem with uncertain \rebut{$\log(A_i)$ values as discussed earlier}.
For the purpose of verification,
$\hat{\mat{C}}$ was initially constructed using a large set of samples
($N$~=~1000) in the input domain. The gradient was estimated using
finite difference, and hence, a total of 20,000 model runs were performed. 
In Figure~\ref{fig:eig_comp}, we illustrate the comparison of
the resulting normalized eigenvalue spectrum \rebut{by plotting $\log(\lambda_i/\lambda_0)$} ($i = 1,\ldots,19$) corresponding to $N$~=~1000 and the same quantity corresponding to
a much smaller set of samples, $n$~=~$\{20,40,80,120\}$.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp_p3}
%   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1}
\caption{A comparison of the normalized eigenvalue spectrum, $\log(\lambda_i/\lambda_0)$
using $n$~=~$\{20,40,80,120\}$ samples with that
obtained using a much larger sample size, $N$~=~1000. 
%Right: Relative $L_2$ 
%norm of the difference between
%the square of individual components ($w_i$'s) of the dominant eigenvector evaluated using 
%$N$=1000 and $n$~=~$\{20,40,80,120\}$ samples.
} 
\label{fig:eig_comp}
\end{center}
\end{figure}
%
We observe that the dominant eigenvalues, $\lambda_1, \ldots, \lambda_4$, 
are approximated 
reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
to improve with the sample size. Since 
$\lambda_1$ is roughly \rebut{an order of magnitude} larger than $\lambda_2$, we expect 
a 1-dimensional active subspace \rebut{to reasonably approximate the uncertainty
in the ignition delay}. 
To further confirm this, we evaluate a relative L$^2$ norm of the difference
 ($\varepsilon_{\text{L}^{2}}^{N-n}$) between the 
squared value of corresponding components of the dominant eigenvector, computed using $N$~=~1000 ($\vec{w}_{1,N}$)
and $n$~=~$\{20,40,80,120\}$ ($\vec{w}_{1,n}$) as follows:
%
\be
\varepsilon_{\text{L}^{2}}^{N-n} = 
\frac{\|\vec{w}_{1,N}^2-\vec{w}_{1,n}^2\|_2}{\|\vec{w}_{1,N}^2\|_2}
\label{eq:accu}
\ee
%
The quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$, was found to be 
$\mathcal{O}(10^{-2})$ in all cases.
Thus, even a small sample size, $n$ = 20, seems to approximate the dominant eigenspace with
reasonable accuracy in this case. 
The iterative strategy therefore offers a significant potential for computational gains.

@@

estimates of the gradient that does not require model evaluations at neighboring points as

@@

strategies for the same amount of computational effort, quantified by the number of model runs, $M$.
The legend provides the number of samples, $N$ and correspondingly, the
model runs, $M$ required for obtaining a converged dominant eigenvector using the perturbation and
regression strategies. Corresponding value of $N$ and $M$ for the test case are also provided.}}

@@

for a test case involving $N$~=~1000 samples using the perturbation approach are also plotted. 

@@

As mentioned earlier, the model output $f(\bm{\xi})$ i.e. the ignition delay in the
H$_2$/O$_2$ reaction in this case, varies
 predominantly in \rebut{a 1-dimensional} active subspace. Hence, 
$f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
the \rebut{1-dimensional active} subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
regarded as the \textit{sufficient summary plot} (SSP), obtained using the perturbation-based and
regression-based gradient estimates are
compared in Figure~\ref{fig:comp_ssp}.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp19D}
\caption{\rebut{An illustrative comparison of the SSPs generated using the 
perturbation and the regression strategies for computing the active subspace.}}
\label{fig:comp_ssp}
\end{center}
\end{figure}
%
\rebut{The dominant eigenvector obtained using perturbation is based on $N$~=~45 samples which requires
$M$~=~900 model runs. For the same amount of computational effort, we can afford $N$~=~900 samples
when using regression. Hence, the SSP from regression is based on 900
points: ($\hat{\mat{W}}_1^\top\bm{\xi}$,$G(\hat{\mat{W}}_1^\top\bm{\xi})$). On the other hand, the SSP
from perturbation is plotted using only 45 points as mentioned earlier.
Nevertheless, the illustrative comparison clearly indicates that the two SSPs
are in excellent agreement. Moreover, it is interesting to note
that the response in ignition
delay based on the considered probability distributions for $\log(A_i)$ although non-linear, can be approximated by
a 1-dimensional active subspace.}

@@

The above comparisons indicate that the gradient of the ignition delay with respect to the
uncertain $A_i$'s is reasonably approximated using both, perturbation-based and regression-based
approaches in this case. Since the regression approach requires fewer model evaluations for
gradient estimation, it might be a preferred approach for this setting.
In the following section, we shift our focus to 
the higher-dimensional H$_2$/O$_2$ reaction kinetics application \rebut{wherein the activation
energies in the rate law as well as initial pressure, temperature, and stoichiometric conditions
are also considered to be uncertain.}

>>>>

In this section, we present an efficient framework for computing the
active subspace. The framework is employed to the H$_2$/O$_2$ reaction kinetics
problem whereby the \rebut{logarithm of the} 
pre-exponent ($A_i$) in the rate law associated with individual reactions provided in
Table~\ref{tab:kinetics} is considered to be uniformly
distributed in the interval, \rebut{$[0.97\log(A_i^\ast), 1.03\log(A_i^\ast)]$};
$A_i^\ast$ is the nominal
estimate provided in~\cite{Yetter:1991}.
Two approaches are explored for estimating the gradient of ignition delay with
respect to \rebut{$\log(A_i)$}: 
a perturbation  approach that involves computation of model
gradients using finite difference in order to construct the matrix $\hat{\mat{C}}$
in~\eqref{eq:chat}, and a regression  approach 
that involves a linear regression fit to the available set of
model evaluations in order to approximate the gradient. 
The active subspace is computed in an iterative manner to avoid
unnecessary model evaluations once converged is established.

@@

As discussed earlier, gradient estimation using finite differences 
requires additional model evaluations at the
neighboring points in the input domain. 
Hence, for $N$ samples in a $d$-dimensional parameter space, $N(d+1)$
model evaluations are needed. On the other hand, gradient estimation
using the regression-based approach involves a series of linear regression
fits to subsets of available evaluations as discussed 
in~\cite{Constantine:2015} (see Algorithm 1.2). Hence, the computational effort is  
reduced by a factor $(d+1)$ when using the regression-based approach.
In other words, for the same amount of computational effort, the regression
approach can afford a sample size that is (d+1) times larger than that in the
case of perturbation approach.
The specific sequence of steps
for computing the active subspace is discussed as follows.

We begin by evaluating the gradient of the model output, $\nabla_{\bm{\xi}}f$,
at an initial set of $n_0$ samples (generated using Monte Carlo sampling)
denoted by $\bm{\xi}_i$, $i$ = $1,\ldots,n_0$.
Using the gradient
evaluations, the matrix, $\hat{\mat{C}}$ is computed. Eigenvalue decomposition
of $\hat{\mat{C}}$ yields an initial estimate of the dominant eigenspace,
$\hat{\mat{W}}_1$ and the set of corresponding eigenvalues, 
$\hat{\mat{\Lambda}}_1$. Note that $\hat{\mat{W}}_1$ is obtained by
partitioning the eigenspace around $\lambda_j$ such that the ratio of
subsequent eigenvalues,
\rebut{$\left(\frac{\lambda_j}{\lambda_{j+1}}\right)\ge\mathcal{O}(10^1)$}.
 At each subsequent iteration, model evaluations are
generated at a new set of $n_k$ samples. The new set of gradient evaluations
are augmented with the available set to re-construct $\hat{\mat{C}}$ followed
by its eigenvalue decomposition. The relative change in the norm of the
difference in squared value of individual components of the dominant
eigenvectors between subsequent iterations is evaluated. The process is
terminated and the resulting eigenspace is considered to have converged once
the maximum relative change at iteration $k$, $\max(\delta \hat{\mat{W}}_{1,j}^{(k)})$
($j$ is used as an index for the eigenvectors),
is smaller than a given tolerance, $\tau$.  A regression fit to
$G(\hat{\mat{W}}_1^\top\bm{\xi})$ is used as a surrogate to characterize and
quantify the uncertainty in the model output. Moreover, the components of the
eigenvectors in the active subspace are used to compute the activity scores, $\bm{\nu}_r(f)$,
which provide an insight into the relative importance of the uncertain inputs.
Note that the index, $r$, corresponds to the number of eigenvectors in
$\hat{\mat{W}}_1$. The sequence of steps as discussed are outlined
in Algorithm~\ref{alg:grad}.

@@

To assess its feasibility and suitability, we implement
Algorithm~\ref{alg:grad} to compute the active subspace for 
the 19-dimensional H$_2$/O$_2$ reaction kinetics
problem by perturbing \rebut{$\log(A_i)$ by 3$\%$ about its nominal
value as discussed earlier}. For the purpose of verification,
$\hat{\mat{C}}$ was initially constructed using a large set of samples
($N$~=~1000) in the input domain. The gradient was estimated using
finite difference, and hence, a total of 20,000 model runs were performed. 
In Figure~\ref{fig:eig_comp}, we illustrate the comparison of
the resulting normalized eigenvalue spectrum \rebut{by plotting $\log(\lambda_i/\lambda_0)$} ($i = 1,\ldots,19$) corresponding to $N$~=~1000 and the same quantity corresponding to
a much smaller set of samples, $n$~=~$\{20,40,80,120\}$.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp_p3}
%   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1}
\caption{A comparison of the normalized eigenvalue spectrum, $\log(\lambda_i/\lambda_0)$
using $n$~=~$\{20,40,80,120\}$ samples with that
obtained using a much larger sample size, $N$~=~1000. 
%Right: Relative $L_2$ 
%norm of the difference between
%the square of individual components ($w_i$'s) of the dominant eigenvector evaluated using 
%$N$=1000 and $n$~=~$\{20,40,80,120\}$ samples.
} 
\label{fig:eig_comp}
\end{center}
\end{figure}
%
We observe that the dominant eigenvalues, $\lambda_1, \ldots, \lambda_4$, 
are approximated 
reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
to improve with the sample size. Since 
$\lambda_1$ is roughly \rebut{an order of magnitude} larger than $\lambda_2$, we expect 
a 1-dimensional active subspace \rebut{to reasonably approximate the uncertainty
in the ignition delay}. 
To further confirm this, we evaluate a relative L$^2$ norm of the difference
 ($\varepsilon_{\text{L}^{2}}^{N-n}$) between the 
squared value of corresponding components of the dominant eigenvector, computed using $N$~=~1000 ($\vec{w}_{1,N}$)
and $n$~=~$\{20,40,80,120\}$ ($\vec{w}_{1,n}$) as follows:
%
\be
\varepsilon_{\text{L}^{2}}^{N-n} = 
\frac{\|\vec{w}_{1,N}^2-\vec{w}_{1,n}^2\|_2}{\|\vec{w}_{1,N}^2\|_2}
\label{eq:accu}
\ee
%
The quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$, was found to be 
$\mathcal{O}(10^{-2})$ in all cases.
Thus, even a small sample size, $n$ = 20, seems to approximate the dominant eigenspace with
reasonable accuracy in this case. 
The iterative strategy therefore offers a significant potential for computational gains.

@@

estimates of the gradient that do not require model evaluations at neighboring points as

@@

strategies for the same amount of computational effort, quantified by the number of model runs, $M$.}}

@@

for a test case involving $N$~=~1000 (and $M$ = 20,000 model runs)
samples using the perturbation approach are also plotted. 

@@

As mentioned earlier, the model output $f(\bm{\xi})$ i.e. the ignition delay in the
H$_2$/O$_2$ reaction in this case, varies
 predominantly in \rebut{a 1-dimensional} active subspace. Hence, 
$f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
the \rebut{1-dimensional active} subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
regarded as the \textit{sufficient summary plot} (SSP), obtained using the perturbation-based and
regression-based gradient estimates are
compared in Figure~\ref{fig:comp_ssp}.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp19D}
\caption{\rebut{An illustrative comparison of the SSPs generated using the 
perturbation and the regression strategies for computing the active subspace.}}
\label{fig:comp_ssp}
\end{center}
\end{figure}
%
\rebut{The dominant eigenvector obtained using perturbation is based on $N$~=~45 samples which requires
$M$~=~900 model runs. For the same amount of computational effort, we can afford $N$~=~900 samples
when using regression. Hence, the SSP from regression is based on 900
points: ($\hat{\mat{W}}_1^\top\bm{\xi}$,~$G(\hat{\mat{W}}_1^\top\bm{\xi})$). On the other hand, the SSP
from perturbation is plotted using only 45 points as mentioned earlier.
Nevertheless, the illustrative comparison clearly indicates that the two SSPs
are in excellent agreement. Moreover, it is interesting to note
that the response in ignition
delay based on the considered probability distributions for $\log(A_i)$ although non-linear, can be approximated by
a 1-dimensional active subspace.}

@@

The above comparisons indicate that the gradient of the ignition delay with respect to the
uncertain $A_i$'s is reasonably approximated using both, perturbation  and regression 
approaches in this case. \rebut{Since both approaches yield consistent results and are
comparable in terms of convergence and accuracy, we could use either for the purpose
of active subspace computation for this setting.}
In the following section, we shift our focus to 
the higher-dimensional H$_2$/O$_2$ reaction kinetics application \rebut{wherein the activation
energies in the rate law as well as initial pressure, temperature, and stoichiometric conditions
are also considered to be uncertain.}
